\newcommand\wgTitle{Add an iota object for simd (and more)}
\newcommand\wgName{Matthias Kretz <m.kretz@gsi.de>}
\newcommand\wgDocumentNumber{P3319R0}
\newcommand\wgGroup{LEWG}
\newcommand\wgTarget{\CC{}26}
%\newcommand\wgAcknowledgements{Daniel Towner and Ruslan Arutyunyan contributed to this paper via discussions / reviews. Thanks also to Jeff Garland for reviewing.}

\usepackage{mymacros}
\usepackage{wg21}
\setcounter{tocdepth}{2} % show sections and subsections in TOC
\hypersetup{bookmarksdepth=5}
\usepackage{changelog}
\usepackage{underscore}
\usepackage{multirow}

\addbibresource{extra.bib}

\newcommand\simd[1][]{\type{ba\-sic\_simd#1}\xspace}
\newcommand\simdT{\type{ba\-sic\_simd\MayBreak<\MayBreak{}T>}\xspace}
\newcommand\valuetype{\type{val\-ue\_type}\xspace}
\newcommand\referencetype{\type{ref\-er\-ence}\xspace}
\newcommand\mask[1][]{\type{ba\-sic\_simd\_mask#1}\xspace}
\newcommand\maskT{\type{ba\-sic\_simd\_mask\MayBreak<\MayBreak{}T>}\xspace}
\newcommand\wglink[1]{\href{https://wg21.link/#1}{#1}}

\newcommand\nativeabi{\UNSP{native-abi}}
\newcommand\deducet{\UNSP{deduce-t}}
\newcommand\simdsizev{\UNSP{simd-size-v}}
\newcommand\simdsizetype{\UNSP{simd-size-type}}
\newcommand\simdselect{\UNSP{simd-select-impl}}
\newcommand\maskelementsize{\UNSP{mask-element-size}}
\newcommand\integerfrom{\UNSP{integer-from}}
\newcommand\constexprwrapperlike{\UNSP{constexpr-wrapper-like}}

\renewcommand{\lst}[1]{Listing~\ref{#1}}
\renewcommand{\sect}[1]{Section~\ref{#1}}
\renewcommand{\ttref}[1]{Tony~Table~\ref{#1}}
\renewcommand{\tabref}[1]{Table~\ref{#1}}

\begin{document}
\selectlanguage{american}
\begin{wgTitlepage}
  There is one important constant in SIMD programming: \code{{0, 1, 2, 3,
  ...}}. In the standard library we have an algorithm called \code{iota} that
  can initialize a range with such values. For \code{simd} we want to have
  simple to spell constants that scale with the SIMD width. This paper proposes
  a simple facility that can be generalized.
\end{wgTitlepage}

\pagestyle{scrheadings}

\input{changelog}
\input{strawpolls}

\section{Motivation}
The 90\% use case for simd generator constructors is a simd with values 0, 1,
2, 3, \ldots{} potentially with scaling and offset applied.
However, often it would be more easier and more readable to use an “iota”
\code{simd} object instead.

\begingroup
  \smaller[1]
\begin{tabular}{p{.45\textwidth}|p{.45\textwidth}}
  generator ctor & iota \\
  \hline
  \begin{lstlisting}
std::simd<int> a([](int i) { return i; };

std::simd<int> b([](int i) { return 2 + 3 * i; };
  \end{lstlisting}
  &
  \begin{lstlisting}
auto a = std::iota_v<std::simd<int>>;

auto b = 2 + 3 * std::iota_v<std::simd<int>>;
  \end{lstlisting}
\end{tabular}
\endgroup

The minimal definition I propose for \simd can look like this:
\medskip\begin{lstlisting}[style=Vc]
  template <class T>
    inline constexpr T
    iota_v;

  template <class T>
    requires(std::is_arithmetic_v<T>)
    inline constexpr T
    iota_v<T> = T();

  template <detail::simd_type T>
    inline constexpr T
    iota_v<T> = T([](auto i) { return static_cast<typename T::value_type>(i); });
\end{lstlisting}


\section{Generalization}
If we define a (\code{constexpr}) variable template \std\code{iota_v<T>} where
\code{T} must be a \simd type, we're simply filling a sequence of values.
We can create such an object for any type with static extent.
This is especially interesting for the degenerate case in SIMD-generic
programming, where \code{T} could \eg be an \code{int}.
A \std\code{iota_v<int>} is nothing other than an object \code{int} with value
\code{0}.
We can easily generalize to \code{iota_v<std::array<T, N>>} and
\code{iota_v<T[N]>}.
And the next step then is to allow any type that
\begin{itemize}
  \item has a static extent,
  \item has a \code{value_type} member type,
  \item can be list-initialized with \code{N} numbers of type
    \code{value_type}, where \code{N} equals the static extent of the type, and
  \item where \code{value_type() + 1} is an constant expression and convertible to \code{value_type}.
\end{itemize}

Consequently you could write
\medskip\begin{lstlisting}[style=Vc]
auto x = std::iota_v<float[5]>;
auto y = std::iota_v<std::array<my_fixed_point, 8>>;
// ...
\end{lstlisting}

\section{Relation to list-initialization of \code{simd}}
If we add a constructor to \simd that enables list-initialization, then many
users might use that in place of a generator constructor.
This leads to code that doesn't scale with the vector width anymore.
Therefore we should provide a simple facility that works better and is more
portable.

\section{Proposed polls}

\wgPoll{We want an iota facility for \simd}{&&&&}

\wgPoll{The iota facility should be generalized to scalars}{&&&&}

\wgPoll{The iota facility should be generalized to any sequence of static extent}{&&&&}

\section{Wording}\label{sec:wording}

TBD after deciding on the preferred solution.

\end{document}
% vim: sw=2 sts=2 ai et tw=0
