\section{Straw Polls}
\subsection{SG1 at Chicago 2013}
\wgPoll{Pursue SIMD/data parallel programming via types?}{1 & 8 & 5 & 0 & 0}

\subsection{SG1 at Urbana 2014}
\begin{itemize}
  \item \wgPoll{SF = ABI via namespace, SA = ABI as template parameter}{0 & 0 & 6 & 11 & 2}
  \item \wgPoll{Apply size promotion to vector operations? SF = \type{shortv} + \type{shortv} = \type{intv}}{1 & 2 & 0 & 6 & 11}
  \item \wgPoll{Apply “sign promotion” to vector operations? SF = \type{ushortv} + \type{shortv} = \type{ushortv}; SA = no mixed signed/unsigned arithmetic}{1 & 5 & 5 & 7 & 2}
\end{itemize}

\subsection{SG1 at Lenexa 2015}
\wgPoll{Make vector types ready for LEWG with arithmetic, compares, write-masking, and math?}{10 & 6 & 2 & 0 & 0}

\subsection{SG1 at JAX 2016}
\begin{itemize}
  \item \wgPoll{Should subscript operator return an lvalue reference?}{0  & 6 & 10 & 2 & 1}
  \item \wgPoll{Should subscript operator return a “smart reference”?}{1  & 7 & 10 & 0 & 0}
  \item \wgPoll{Specify simd width using ABI tag, with a special template tag for fixed size.}{3 & 7 & 0 & 0 & 1}
  \item \wgPoll{Specify simd width using <T, N, abi>, where abi is not specified by the user.}{1 & 2 & 5 & 2 & 1}
\end{itemize}

\subsection{SG1 at Oulu 2016}
\begin{itemize}
  \item \wgPoll{Keep \code{native_handle} in the wording (dropping the ampersand in the return type)?}{0 & 6 & 3 & 3 & 0}
  \item \wgPoll{Should the interface provide a way to specify a number for over-alignment?}{2 & 6 & 5 & 0 & 0}
  \item \wgPoll{Should loads and stores have a default load/store flag?}{0 & 0 & 7 & 4 & 1}
\end{itemize}

\subsection{LEWG at Issaquah 2016}
\begin{itemize}
  \item \wgPoll{Unary minus on unsigned simd should be ill formed}{0 & 5 & 6 & 0 & 3}
  \item \wgUnanimous{Reductions only as free functions}
  \item \wgUnanimous{Jens should work with the author and return with an updated paper}
\end{itemize}

\subsection{LEWG at Kona 2017}
\begin{itemize}
  \item \wgUnanimous{Want \code{operator<<(signed)} to work except where it's undefined for the underlying integer?}

  \item \wgPoll{Should there be overloads for \code{where} and (\mask and \simd) reductions with \bool/builtin types in place of \mask/\simd?}{1&5&6&1&0}

  \item \wgUnanimous{Should there be a named “widen” function that widens the element type \type T from (e.g.) \intt to \type{long}, but rejects \intt to \short?
  The number of elements is not changed.}
  cf. \autoref{sec:widen}

\item \wgUnanimous{Should there be a named “concat” functions that concatenates several \simd{}s with the same element type, but potentially different length?\\
  The latter two operations are currently lumped together in \code{simd_cast}.\\
  The widen / concat questions, I think, comes down to whether we want the current \code{simd_cast} or not.}
  cf. \autoref{sec:split and concat}

\item \wgUnanimous{We still have a few open ends on implicit conversions (\url{https://github.com/mattkretz/wg21-papers/issues/26}, \url{https://github.com/mattkretz/wg21-papers/issues/3}).
  The current paper is rather strict, there may be room for more implicit conversions without introducing safety problems.
Should we postpone any changes in this area to after TS feedback?}

\item \noindent Poll: Keep or drop \code{simd_abi::max_fixed_size}? (cf. \url{https://github.com/mattkretz/wg21-papers/issues/38})\\
  \hspace*{2em} Authors will come back with a proposal\par\medskip
  cf. \autoref{sec:maxfixedsize}

\item \wgUnanimous{Add a list of possible names for “simd” and “where” to the paper.}
  cf. \autoref{sec:Naming}
\end{itemize}

\subsection{LEWG at Toronto 2017}
\begin{itemize}
  \item \wgUnanimous{Use Option 3 (support both element type and \simd type as cast template argument, instead of just either one of them) for \code{static_simd_cast} and \code{simd_cast}.}

  \item \wgUnanimous{Use \code{abi_for_size_t} instead of unconditionally using \code{fixed_size} for split and concat.
  Revisit the question after TS feedback.}

  \item \wgPoll{Name it \code{datapar} (in favor) or \code{simd} (against)?}
    {1&4&1&5&4}
    We want TS feedback on the name, i.e. whether anyone is confused about the applicability of the feature because of the name.

  \item \wgPoll{Name \code{mask} (in favor) or \code{simdmask} (against)?}
    {0&1&2&5&7}

  \item \wgPoll{Name \code{simd_mask} (in favor) or \code{simdmask} (against)?}
    {4&6&3&1&0}

  \item \wgPoll{\code{where} (in favor) vs. \code{masked} (against)?}
    {5&1&3&2&3}
    No consensus for change, revisit after TS feedback.

  \item \wgVote{Vote for your preference for a name for the load/store functions}
    {r|r|r|r}
    {
      \code{load}&\code{memload}&\code{load_from}&\code{copy_from}\\
      4&2&6&7
    }

  \item \wgPoll{\code{load_from} (in favor) vs. \code{copy_from} (against)?}
    {2&3&3&0&3}
    No consensus to override the author's preference (\code{copy_from}).

  \item \wgUnanimous{Keep \code{simd_mask} reductions as free functions with the current names.
  Consider follow-up papers if any names need to be changed (e.g. for consistency).}

  \item \wgPoll{Forward to LWG for TS}{11&2&0&0&0}
\end{itemize}

% vim: sw=2 sts=2 ai et tw=0
