\section{Wording}

The following is a draft targeting inclusion into the Parallelism TS 2.
It defines a basic set of data-parallel types and operations.

\newcommand\clause{Clause\xspace}
\newcommand\width{width\xspace}

\begin{wgText}
  \setcounter{WGClause}{7}
  \wgSection{Data-Parallel Types}{simd.types}
  \wgSubsection{General}{simd.general}
  \pnum
  The data-parallel library consists of data-parallel types and operations on these types.
  A data-parallel type consists of elements of an underlying arithmetic type, called the \emph{element type}.
  The number of elements is a constant for each data-parallel type and called the \emph{\width} of that type.

  \pnum
  Throughout this \clause, the term \emph{data-parallel type} refers to all \emph{supported} \ref{sec:simd.overview} specializations of the \simd and \mask class templates.
  A \emph{data-parallel object} is an object of \emph{data-parallel type}.

  \pnum
  An \emph{element-wise operation} applies a specified operation to the elements of one or more data-parallel objects.
  Each such application is unsequenced with respect to the others.
  A \emph{unary element-wise operation} is an element-wise operation that applies a unary operation to each element of a data-parallel object.
  A \emph{binary element-wise operation} is an element-wise operation that applies a binary operation to corresponding elements of two data-parallel objects.

  \pnum Throughout this \clause, the set of \emph{vectorizable types} for a data-parallel type comprises all cv-unqualified arithmetic types other than \bool.

  \pnum \label{cl:intent-note}\wgNote{
      The intent is to support acceleration through data-parallel execution resources, such as SIMD registers and instructions or execution units driven by a common instruction decoder.
      If such execution resources are unavailable, the interfaces support a transparent fallback to sequential execution.
  }

  \input{wording-synopsis}
  \input{wording-simd}
  \input{wording-mask}
\end{wgText}

% vim: tw=0
