\newcommand\wgTitle{Floating-Point Excess Precision}
\newcommand\wgName{Matthias Kretz <m.kretz@gsi.de>}
\newcommand\wgDocumentNumber{DFPEPR0}
\newcommand\wgGroup{SG6, EWG, CWG}
\newcommand\wgTarget{\CC{}26}
%\newcommand\wgAcknowledgements{ }

\usepackage{mymacros}
\usepackage{wg21}
%\setcounter{tocdepth}{2} % show sections and subsections in TOC
%\hypersetup{bookmarksdepth=5}
\usepackage{changelog}
\usepackage{underscore}
%\usepackage{comment}

\addbibresource{extra.bib}

\renewcommand{\lst}[1]{Listing~\ref{#1}}
\renewcommand{\sect}[1]{Section~\ref{#1}}
\renewcommand{\ttref}[1]{Tony~Table~\ref{#1}}

\begin{document}
\selectlanguage{american}
\begin{wgTitlepage}
  CWG2752 asks about the design intent for handling excess-precision
  floating-point computations in \CC{}.
  The main question we need answered is whether literals must be rounded or can
  be stored with excess precision.
  But as soon as we start to touch the existing wording, we need to make sure
  it still matches the design intent.
  This paper is trying to establish a consensus on design intent by listing
  potential optimizations a compiler could do and deciding whether they should
  be conforming or not.
\end{wgTitlepage}

\pagestyle{scrheadings}

%\input{changelog}
%\input{strawpolls}

\section{Introduction}

\section{Possible Implications of Excess Precision}

\begin{itemize}
  \item x87 FPU usage
    \begin{itemize}
      \item At what points is intermediate rounding required?
      \item Do parenthesis have any meaning?
    \end{itemize}

  \item fp contraction into FMAs: \code{a * b + c} translated to \code{fma(a, b, c)}
    \begin{itemize}
      \item Should \code{(a * b) + c} ever make a difference?
        I.e. ensure rounding after multiplication?
    \end{itemize}

  \item Given two floating-point values \code{a} and \code{b} where \code{std::isfinite(b)} is statically known to be \code{true}:
    \begin{itemize}
      \item Is the optimizer allowed to replace \code{a + b - b} with \code{a}?
      \item Is the optimizer allowed to replace \code{b - b} with \code{0}?
      \item Is the optimizer allowed to replace \code{a + a + a} with \code{3 * a}?
      \item Do parenthesis have additional meaning, inhibiting any of the above replacements?
        (Is \code{a + b - b} different from \code{(a + b) - b}?)
      \item Consider the case where \code{std::isfinite(a)} and \code{std::isinf(a + b)}.
      \item Consider negative zero.
    \end{itemize}

  \item excess precision literals
\end{itemize}

\section{Interaction with library wording}

\begin{itemize}
  \item \code{FLT_EVAL_METHOD}?
\end{itemize}

\end{document}
% vim: sw=2 sts=2 ai et tw=0
