\newcommand\wgTitle{value-preserving consteval broadcast to \code{simd::vec}}
\newcommand\wgName{Matthias Kretz <m.kretz@gsi.de>}
\newcommand\wgDocumentNumber{P4012R0}
\newcommand\wgGroup{LEWG, LWG}
\newcommand\wgTarget{\CC{}26}
%\newcommand\wgAcknowledgements{Daniel Towner and Ruslan Arutyunyan contributed to this paper via discussions / reviews. Thanks also to Jeff Garland for reviewing.}

\usepackage{mymacros}
\usepackage{wg21}
\setcounter{tocdepth}{2} % show sections and subsections in TOC
\hypersetup{bookmarksdepth=5}
\usepackage{changelog}
\usepackage{underscore}
\usepackage{multirow}
\usepackage{vectorpictures}

\addbibresource{extra.bib}

\newcommand\wglink[1]{\href{https://wg21.link/#1}{#1}}

\renewcommand{\lst}[1]{Listing~\ref{#1}}
\renewcommand{\sect}[1]{Section~\ref{#1}}
\renewcommand{\ttref}[1]{Tony~Table~\ref{#1}}
\renewcommand{\tabref}[1]{Table~\ref{#1}}

\begin{document}
\selectlanguage{american}
\begin{wgTitlepage}
  The broadcast constructor in the Parallelism 2 TS allowed construction from (\code{unsigned})
  \code{int}, allowing e.g.\ \code{vec<float>() + 1}, which is ill-formed in the CD\@.
  This breaks existing code that gets ported from the TS to \code{std::simd}.
  The design intent behind \code{std::simd} was for this to work.
  However, the understanding in LEWG appeared to be that we can't get this right without constexpr
  function arguments getting added to the language.
  This paper shows that a \code{consteval} constructor overload together with \code{constexpr}
  exceptions can resolve the issue for \CC{}26 and is a better solution than constexpr function
  arguments would be.
\end{wgTitlepage}

\pagestyle{scrheadings}

\input{changelog}
\input{strawpolls}

\section{The Proposal}
\begin{enumerate}
  \item Add a \code{consteval} broadcast constructor overload that is constrained to a minimal set
    of not-value-preserving types.
  \item Reword the constraint of the existing broadcast constructor for correct partial ordering
    with the new \code{consteval} overload.
  \item The presence of the new overload effectively reverts a small part of “Issue 1” (Section 3)
    of \parencite{P3430R3}.
\end{enumerate}

The consequence of this change is summarized in~\ttref{tt:effect}.
\begin{tonytable}{Effect of this paper}\label{tt:effect}
  \begin{lstlisting}
  int n = 1;
  // ------ no change: --------
  simd::vec<float> x = 1.f;
  // x = 0x5EAF00D; // ill-formed
  // x = n; // ill-formed
  x = static_cast<float>(n); // OK
  static_assert(constructible_from<V, int>);

  // ------ different: --------
  // x = 1; // ill-formed
  x = static_cast<simd::vec<float>>(n); // OK
  // pow(x, 3); // ill-formed
  static_assert(not convertible_to<int, V>);
  // common_type_t<V, int> y = x; // ill-formed
  \end{lstlisting}
  &
  \begin{lstlisting}
  int n = 1;
  // ------ no change: --------
  simd::vec<float> x = 1.f;
  // x = 0x5EAF00D; // ill-formed
  // x = n; // ill-formed
  x = static_cast<float>(n); // OK
  static_assert(constructible_from<V, int>);

  // ------ different: --------
  x = 1; // OK
  // x = static_cast<simd::vec<float>>(n); // ill-formed
  pow(x, 3); // OK
  static_assert(convertible_to<int, V>); // opposite
  common_type_t<V, int> y = x; // V
  \end{lstlisting}
\end{tonytable}%
To understand why the concepts work out that way, consider that concepts (and traits) basically only
work at the type level (and overload resolution rules).
Whether an immediate function is called with an argument that is not a constant expression is simply
not part of the consideration.
The types match, overload resolution does not discriminate on \code{consteval} and so the following
example comes out seemingly wrong:
\medskip\begin{lstlisting}[style=Vc]
struct X { consteval X(int) {} };
void f(X);
int g();

static_assert(requires { f(g()); }); // OK
void test()            { f(g()); }   // error: call to consteval 'X::X'
\end{lstlisting}
The requires expression says \code{f(g())} is fine, but when we actually use it, it is ill-formed.
This is the same “feature” which leads to mischaracterization of \code{constructible_from} and
\code{convertible_to} with the \code{basic_vec} \code{consteval} broadcast constructor.

\section{The Concerns}
It quickly became apparent that LWG would not feel comfortable voting this paper to plenary.
While I had considered all the concerns that were brought up, I did not feel like LEWG had taken the
time in Kona to understand these concerns and think it all through.
Under these circumstances I don't want to force a contested vote in plenary but rather make sure
this is well understood.

The concerns raised in LWG:
\begin{enumerate}
  \item Novel (no other type in the standard library does this\footnote{\code{format} strings are
    very close, though: \url{https://compiler-explorer.com/z/7z8Y5ooj3}}).
  \item Constrained code can be ill-formed, because the traits/concepts lie.
  \item Immediate-escalation could force more code to be immediate than was intended by the user.
  \item Errors after immediate-escalation are potentially harder to understand.
\end{enumerate}

\section{Motivation}

It is very common in floating-point code to simply write e.g. \code{* 2} rather than \code{* 2.f}
when multiplying a \code{float} with a constant:
\medskip\begin{lstlisting}[style=Vc]
float f(float x) { return x * 2; } // converts 2 to float (at compile time)

float g(float x) { return x * 2.; } // converts x to double (at run time)

float h(float x) { return x * 2.f; } // no conversions
\end{lstlisting}

More importantly, using \code{* 2} works reliably in generic code, where the type of \code{x} could
be any arithmetic type.

Since this is so common, \code{std::experimental::simd<T>} made an exception for \code{int} in the
broadcast constructor to not require value-preserving conversions.
Consequently, the TS behavior is:
\medskip\begin{lstlisting}[style=Vc]
using floatv = std::experimental::native_simd<float>;

floatv f(floatv x) { return x * 2; } // converts 2 to float and broadcasts (at
                                     //                               compile time)
floatv g(floatv x) { return x * 2.; } // ill-formed

floatv h(floatv x) { return x * 2.f; } // broadcasts 2.f to floatv
\end{lstlisting}

When porting existing code written against the TS to \CC{}26, the first step is to adjust the types:
\medskip\begin{lstlisting}
using floatv = std::@\wgChange{experimental::native_simd}{simd::vec}@<float>;
\end{lstlisting}
Except for uses of \code{std::experimental::where}, which need to be refactored to use
\code{simd::\MayBreak{}select}, the remaining code should work.
The one place where it doesn't work is code such as in function \code{f}, where \code{2} needs to be
replaced:
\medskip\begin{lstlisting}[style=Vc]
floatv f(floatv x) { return x * @\wgChange{2}{std::cw<2>}@; }
\end{lstlisting}

Since we don't have \code{constexpr} function arguments in the language, \code{std::simd} works
around it by recognizing \exposid{integral-constant-like} / \code{constant-wrapper-like} types, that
encode a \emph{value} into a type.
This, however, comes at a compile-time cost.
Every different value leads to a template specialization of both \code{constant_wrapper} and a
\code{basic_vec} broadcast constructor (with it's helper types/concepts to determine whether the
specialization is allowed).
Consequently, for \code{vec<float>}, I would recommend to always use an \code{f} suffix rather than
\code{std::cw}.

But that solution is fairly limited, since we don't have literals for 8-bit and 16-bit integers in
the language.
A function template like
\medskip\begin{lstlisting}[style=Vc]
template <simd_integral V>
V f(V x) {
  return x + 1; // ill-formed for V::value_type = (u)int8_t, (un)int16_t, and uint32_t
}
\end{lstlisting}
needs to use \code{x + V(1)}\footnote{explicit conversion to \code{basic_vec} allows conversions
that are not value-preserving}.
A clever user might write \lstinline@x + '\1'@ instead.
But that fails for the \code{char} type with different signedness.

Consequently, users would need to get used to writing explicit conversions for the constants they
use in \code{std::simd} code.
That's not only verbose and ugly, it is also error-prone.
Whenever we coerce our users into writing explicit conversions, then value-changing conversions
cannot be diagnosed as erroneous anymore.
An explicit \code{static_cast<uint64_t>(-1)} means \code{0xffff'ffff'ffff'ffff}, whereas
\code{uint64_t x = -1} could have been intended to mean \code{0x0000'0000'ffff'ffff} or is a result
of a logic flaw in the code.
E.g., GCC's \code{-Wsign-conversion} diagnoses the latter, but not the former\footnote{And that's
useful, because the former says “I'm intentionally doing this conversion, no need to warn.”}.

If, with \CC{}26, our users are starting to explicitly convert their \code{int} constants to
\code{basic_vec}, then the interface of \code{basic_vec} is at least in part guilty for introducing
harder to find bugs.

\ttref{tt:add-an-offset} presents an example of the solution\footnote{I got bitten by this in my
\code{std::simd} unit tests}.
Note that the code on the left will never warn about the value-changing conversion, even with all
conversion related warnings enabled.
This is due to the explicit conversion, which is telling the compiler “I know what I'm doing; no
need to warn me about it”.
\begin{tonytable}{Add an offset}\label{tt:add-an-offset}
  \begin{lstlisting}
template <simd_floating_point V>
V f(V x) {
  return x + V(0x5EAF00D);
}

f(vec<double>()); // OK

// compiles but adds 99282960 instead of 99282957
f(vec<float>());

// compiles but adds infinity instead of 99282957
f(vec<std::float16_t>());
  \end{lstlisting}
  &
  \begin{lstlisting}
template <simd_floating_point V>
V f(V x) {
  return x + 0x5EAF00D;
}

f(vec<double>()); // OK

// ill-formed: value-changing conversion
f(vec<float>());

// ill-formed: value-changing conversion
f(vec<std::float16_t>());
  \end{lstlisting}
\end{tonytable}%

A safer implementation of the code on the left side of \ttref{tt:add-an-offset} (without this paper)
would have been to write \code{x + std::cw<0x5EAF00D>} instead.
Then the value-changing conversion would have resulted in a constraint failure on the broadcast
constructor.
However, \code{V(0x5EAF00D)} is shorter and needs fewer template instantations.
I expect most users (including myself) will/do not use \code{std::cw} all over the place.

\section{Design space}
In the design review of P1928 of this issue of the broadcast constructor it was overlooked (and
never discussed) that a \code{consteval} overload of the broadcast constructor could solve this
problem.
Before \code{constexpr} exceptions, we would have worded it to be ill-formed (by unspecified means)
if the value changes on conversion to the \code{basic_vec}'s value-type.
Now that we have \code{constexpr} exceptions, we can specify a \code{consteval} broadcast overload
that throws on value-changing conversion.
If the caller cares, the exception can even be handled at compile time.
(I believe it should not throw in \CC{}26, for a minimal change to the WD this late in the \CC{}26
cycle.)

Ordering the overloads for overload resolution is tricky, which is another reason why we should
consider this issue before \CC{}26 ships and potentially take action even if we don't add a
\code{consteval} overload.
Overload resolution does not take \code{consteval} into account.
The process of finding candidate functions%
\href{https://eel.is/c++draft/over.match#funcs.general-8}{\iref{over.match.funcs.general}}, however,
does remove explicit constructors from the candidate set if the
context does not allow the explicit constructor to be called.

\subsection{potentially-convertible-to}
R0 proposed to allow any conversion from arithmetic type \code{U}, that satisfies
\code{convertible_to<value_type>} and does not satisfy
\code{\exposid{value-preserving-convertible-to}<value_type>} via the \code{consteval} broadcast
constructor.
This was too broad, since it would lead to \code{vec<\MayBreak{}float>() + 1.5} being valid
(of type \code{vec<float>}).
While technically not wrong (no loss on conversion from \code{1.5}), it is too surprising that an
operation involving a \code{double} operand is evaluated in single precision.

Therefore, R1 of this paper tightens the constraints for the \code{consteval} broadcast to producing
a less surprising common type.
If the given constant is of arithmetic type \code{T}, then we now require \code{common_type_t<T,
value_type>} to be \code{value_type}.
Since \code{common_type_t<\MayBreak{}double, float>} is \code{double}, the expression
\code{vec<float>() + 1.5} becomes ill-formed.
However, this rule alone still breaks the case of \code{vec<short>() + 1}, which the user cannot
changed to use a \code{short} literal (because we don't have one).
The TS made an explicit exception for \code{int} and \code{unsigned int}\footnote{%
  The TS broadcast constructor has a constraint “[\ldots], or \code{From} is \code{int}, or
  \code{From} is \code{unsigned int} and \code{value_type} is an unsigned integral type.”
}, which is what we still
need for integer types (with lower rank than \code{int}).

So the final \exposid{potentially-convertible-to} constraint looks like this:
\medskip\begin{lstlisting}[style=Vc]
template <typename From, typename To>
  concept @\exposid{potentially-convertible-to}@ = is_arithmetic_v<From>
    && convertible_to<From, To> && !@\exposid{value-preserving-convertible-to}@<From, To>
    && (is_same_v<common_type_t<From, To>, To>
      || (is_same_v<From, int> && is_integral_v<To>)
      || (is_same_v<From, unsigned> && unsigned_integral<To>));
\end{lstlisting}

\subsection{Status quo}
The following code shows the properties of the current broadcast constructor.
See Appendix \ref{sec:really-convertible} for the definition of the \code{really_convertible_to}
concept.

\bigskip
\begin{lstlisting}[style=Vc]
using V = simd::vec<float>;

template <typename T> struct X { explicit operator T() const; };

template <typename... Ts>
concept has_common_type = requires { typename std::common_type_t<Ts...>; };

static_assert(not   std::convertible_to<X<float>, V>);
static_assert(      std::convertible_to<float, V>);
static_assert(      std::convertible_to<short, V>);
static_assert(    really_convertible_to<short, V>);
static_assert(not   std::convertible_to<int, V>);
static_assert(not really_convertible_to<int, V>);

static_assert(    std::constructible_from<V, X<float>>);
static_assert(not std::constructible_from<V, X<short>>);
static_assert(    std::constructible_from<V, double>);
static_assert(    std::constructible_from<V, float>);
static_assert(    std::constructible_from<V, short>);
static_assert(    std::constructible_from<V, int>);

static_assert(not has_common_type<V, double>);
static_assert(not has_common_type<V, int>);

V f(int n, short m, std::reference_wrapper<int> l, std::reference_wrapper<float> f)
{
  V x = '\1'; // OK
  x = 1;      // ill-formed
  x = 0x5EAF00D; // ill-formed
  x = V(n);   // OK
  x = m;      // OK
  x = l;      // OK (because convertible_to<decltype(l), float> is true)
  x = f;      // OK
  x = 1.1;    // ill-formed
  x = V(1.1); // OK
  x = X<float>(); // ill-formed: no match for operator= (no known conversion […])
  x = float(X<float>()); // OK (obvious)
  x = V(X<float>()); // OK
}
\end{lstlisting}

\subsection{The proposed solution}\label{sec:consteval}
The new \code{consteval} constructor is more constrained than the existing constructor (using
subsumption).
Thus, the \code{consteval} constructor is always chosen if the conversion of the given (arithmetic)
type to \code{value_type} is \exposid{potentially-convertible-to}.
Otherwise, the \code{constexpr} overload is used.
As a consequence, some explicit conversions from arithmetic types to \code{basic_vec} do not work
anymore.

Sketch:
\medskip\begin{lstlisting}[style=Vc]
template <class From, class To>
  concept @\exposconcept{simd-consteval-broadcast-arg}@ =
    @\exposconcept{explicitly-convertible-to}@<From, To> && @\exposconcept{potentially-convertible-to}@<remove_cvref_t<From>, To>;

template <class T>
class basic_vec
{
public:
  template <@\exposconcept{explicitly-convertible-to}@<T> U>
    constexpr explicit(@\seebelow@) basic_vec(U&&); // #1
  template <@\exposconcept{simd-consteval-broadcast-arg}@<T> U>
    consteval basic_vec(U&&); // #2
};
\end{lstlisting}

Now, every explicit call to the broadcast constructor with a type \code{U} that satisfies
\code{\exposid{poten\-tial\-ly-convertible-to}<T>} is equivalent to an implicit conversion, since the
\code{consteval} overload is viable and more constrained.
Every type with a value-preserving conversion to \code{T} will select \code{\#1} (because of the
constraint on \code{\#2}).
Every non-arithmetic type (notably, user-defined types with conversion operator to some arithmetic
type) will continue to work as today, since \code{\#2} is not viable.

\bigskip
\begin{lstlisting}[style=Vc]
static_assert(not   std::convertible_to<X<float>, V>); // equal to status quo / different to above
static_assert(      std::convertible_to<float, V>);
static_assert(      std::convertible_to<short, V>);
static_assert(    really_convertible_to<short, V>);
static_assert(      std::convertible_to<int, V>);      // different to status quo / equal to above
static_assert(not really_convertible_to<int, V>);

static_assert(    std::constructible_from<V, X<float>>);
static_assert(not std::constructible_from<V, X<short>>);
static_assert(    std::constructible_from<V, double>);
static_assert(    std::constructible_from<V, float>);
static_assert(    std::constructible_from<V, short>);
static_assert(    std::constructible_from<V, int>);

static_assert(not has_common_type<V, double>);
static_assert(    has_common_type<V, int>); // it's vec<float>

V f(int n, short m, std::reference_wrapper<int> l, std::reference_wrapper<float> f)
{
  V x = '\1'; // OK
  x = 1;      // OK (different to status quo / equal to above)
  x = 0x5EAF00D; // ill-formed
  x = V(n);   // ill-formed (different to both)
  x = m;      // OK
  x = V(l);   // OK
  x = f;      // OK
  x = 1.1;    // ill-formed
  x = V(1.1); // OK
  x = X<float>(); // ill-formed: no match for operator= (no known conversion […])
  x = float(X<float>()); // OK (obvious)
  x = V(X<float>()); // OK
}
\end{lstlisting}

\subsection{How to handle bad value-preserving casts}\label{sec:exceptions}

The \code{consteval} broadcast overload needs to be ill-formed if the argument value cannot be
converted to the value type without changing the value.
This can be achieved via the mechanism used in \iref{simd.bit} for \code{bit_ceil}.
The constructor would spell out a precondition followed by \remarks An expression that violates
the precondition in the \expects element is not a core constant expression\iref{expr.const}.

The alternative that was mentioned before is to throw an exception (at compile time).
Since in virtually all cases, such an exception would not be caught at compile time, the program
becomes ill-formed.
The ability to catch the exception allowed me to hack up a \code{really_convertible_to} concept.
But otherwise, the utility of using an exception here seems fairly limited.
The main reason for using an exception is better diagnostics on ill-formed programs.
If we decide to add the \code{consteval} constructor for \CC{}26, then we might want to delay the
new exception type for \CC{}29, though.

\section{Differences}

\newcommand\good{\color[rgb]{0,.4,.2}\ding{52} }
\newcommand\bad{\color[rgb]{.5,0,0}\ding{56} }

Differences between the status quo and the two alternatives above:\\
\begin{tabular}{l|rrr}
  & status quo & \sect{sec:consteval} \\ \hline
  \code{convertible_to<int, V>}          & \false & \true \\
  \code{common_type_t<V, int>}           & \bad   & \code{V} \\
  \code{x = 1;}                          & \bad   & \good \\
  \code{x = V(n);}                       & \good  & \bad \\
\end{tabular}

\medskip
Note that while some values of constant expressions of type \code{int} are convertible to
\code{vec<float>}, it is not true in general that \code{int} is convertible to \code{vec<float>}.

\section{Implementation experience}
Both solutions (and a lot more variants that were discarded) have been implemented and tested in my
implementation.
Several days (if not weeks) of exploration and testing went into this paper.
I implemented the \code{consteval} overloads for a complete set of vectorizable types with an
ability to select between the different behaviors discussed in \parencite{P3844R1}.
A representative set of [simd.math] is implemented.

\section{Recommendation}

My recommendation continues to be to add the \code{consteval} broadcast overload for \CC{}26.
However, we do have the option of opening up the design space for doing something in \CC{}29.
I acknowledge that the concerns are troubling issues.
Maybe time is all we need to get comfortable.
Maybe something else can be done that we don't think of yet.
Therefore, to keep the design space open for \CC{}29, we can remove explicit conversions on
broadcast.
Explicit conversion on broadcast is not an important feature to have, in my opinion, and removing it
gives us the necessary breathing room for \CC{}29.

\input{wording}
\input{wording_alt}

\appendix
\let\appendix=\relax
\section{\code{really_convertible_to} definition}\label{sec:really-convertible}
\medskip\begin{lstlisting}[style=Vc]
template <typename To, typename From>
  consteval bool converting_limits_throws()
  {
    try {
      using L = std::numeric_limits<From>;
      [[maybe_unused]] To x = L::max();
      x = L::min();
      x = L::lowest();
    } catch(...) {
      return true;
    }
    return false;
  }

template <typename From, typename To>
  concept really_convertible_to = std::convertible_to<From, To>
                                    and not converting_limits_throws<To, From>();
\end{lstlisting}


\end{document}
% vim: sw=2 sts=2 ai et tw=100
