\section{Changelog}
\begin{revision}
  \item Update to apply against \CC{}17 wording.
  \item Removed executors discussion because the executors design has not left SG1 yet.
  \item Updated example code to reflect changes in P0214.
\end{revision}

\begin{revision}
  \item Updated code to match \cite{N4744}.
  \item Fixed a bug in the \code{for_each} example implementation.
  \item Improved \code{iota} and \code{for_each} example implementations with constexpr-if.
  \item Discuss impact on all algorithms.
\end{revision}

\begin{revision}
  \item Discuss ABI tag of \code{std::generate} callables.
  \item Add a Tony Table.
  \item Note that \code{remove} and \code{remove_copy} are implicitly vectorizable.
\end{revision}

\begin{revision}
  \item Add heap algorithms to list of algorithms that may benefit from new execution policy.
  \item Discuss the 3 options for \code{generate}.
  \item Settle on \code{simd_mask} predicates.
\end{revision}

\begin{revision}
  \todo Discuss algorithms in \code{<numeric>}. (transform_reduce is a bit more involved because of different value_types)
  \todo Mention \code{<memory>} algorithms (they don't call user-defined code).
  \todo Add pointer to P0214 for motivation.
  \todo Fix mention of \code{swap_ranges}, which was \code{swap} before.
  \todo Add C++20's \code{shift_left} and \code{shift_right}.
  \todo \code{make_heap}, \code{push_heap}, \code{pop_heap}, and \code{sort_heap} have no execution policy overload. Do fix.
  \todo Discuss permission to fall back to sequential execution (i.e. T vs. simd<T, scalar>).
\end{revision}

  %\todo Discuss that the “drop \code{generate}” option is really unfortunate wrt. random number usage.
  %\todo Discuss concept requirements.
