\newcommand\wgTitle{Integrating datapar with parallel algorithms and executors}
\newcommand\wgName{Matthias Kretz <m.kretz@gsi.de>}
\newcommand\wgDocumentNumber{D0350R1}
\newcommand\wgGroup{LEWG}
\newcommand\wgAcknowledgements{
  This work was supported by GSI Helmholtzzentrum für Schwerionenforschung
  and the Hessian LOEWE initiative through the Helmholtz International Center
  for FAIR (HIC for FAIR).
}

\usepackage{vectorpictures}
\usepackage{typenames}
\usepackage{mymacros}
\usepackage{wg21}
\undef\BreakableUnderscore
\usepackage{underscore}

\addbibresource{extra.bib}

\newcommand\datapar[1][]{\type{datapar#1}\xspace}
\newcommand\valuetype{\type{value\textunderscore{}type}\xspace}
\newcommand\simdcast{\code{datapar\textunderscore{}cast}\xspace}
\newcommand\mask[1][]{\type{mask#1}\xspace}
\newcommand\fixedsizeN{\type{datapar\textunderscore{}abi::fixed\textunderscore{}size<N>}\xspace}
\newcommand\fixedsizescoped{\type{datapar\textunderscore{}abi::fixed\textunderscore{}size}\xspace}
\newcommand\fixedsize{\type{fixed\textunderscore{}size}\xspace}
\newcommand\dataparEP{\code{execution::}\type{datapar}\xspace}
\DeclareRobustCommand\dataparEPT{\code{execution::}\type{datapar\textunderscore{}policy}\xspace}
\newcommand\seqEP{\code{std::execution::}\type{seq}\xspace}

\newcommand\foralli[1][]{for all \code i $\in$ \code{[0, #1size())}\xspace}

\begin{document}
\selectlanguage{american}
\begin{wgTitlepage}
  This paper discusses a new execution policy for integrating datapar with \emph{parallel algorithms}.
\end{wgTitlepage}

\pagestyle{scrheadings}
\addtocounter{section}{-1}
\section{Remarks}
\begin{itemize}
  \item This documents talks about “vector” types/objects.
    In general this will not refer to the \std\type{vector} class template.
    References to the container type will explicitly call out the \code{std} prefix to avoid confusion.
  \item \citep{P0214R4} is the last paper on \datapar.
\end{itemize}

\section{Changelog}
\subsection{Changes from R0}
Previous revision: \parencite{P0350R0}.
\begin{itemize}
  \item Update to apply against \CC{}17 wording.
  \item Removed executors discussion because the design has not left SG1 yet.
  \item Updated example code to reflect changes in P0214.
\end{itemize}

\section{Straw Polls}
\subsection{SG1 at Oulu}
\wgPoll{Ship it to LEWG?}{6 & 6 & 2 & 0 & 0}

\input{intro}
\input{algorithms}

\end{document}
% vim: sw=2 sts=2 ai et tw=0 spell
