\newcommand\wgTitle{\code{constant\_wrapper} should unwrap on call and subscript}
\newcommand\wgName{Matthias Kretz <m.kretz@gsi.de>\\
  Tomasz Kamiński <tomaszkam@gmail.com>}
\newcommand\wgDocumentNumber{P3978R2}
\newcommand\wgGroup{LEWG}
\newcommand\wgTarget{\CC{}26}
\newcommand\wgAcknowledgements{Thanks to Barry Revzin for helpful feedback.}

\usepackage{mymacros}
\usepackage{wg21}
\setcounter{tocdepth}{2} % show sections and subsections in TOC
\hypersetup{bookmarksdepth=5}
\usepackage{changelog}
\usepackage{underscore}
\usepackage{multirow}

\addbibresource{extra.bib}

\newcommand\wglink[1]{\href{https://wg21.link/#1}{#1}}

\renewcommand{\lst}[1]{Listing~\ref{#1}}
\renewcommand{\sect}[1]{Section~\ref{#1}}
\renewcommand{\ttref}[1]{Tony~Table~\ref{#1}}
\renewcommand{\tabref}[1]{Table~\ref{#1}}

\begin{document}
\selectlanguage{american}
\begin{wgTitlepage}
  This paper proposes unwrapping overloads of \code{operator()} and \code{operator[]} to
  \code{constant_wrapper} while making them all \code{static}.
  We also change the call operator to always use \placeholder{INVOKE}.
\end{wgTitlepage}

\pagestyle{scrheadings}

\input{changelog}
\input{strawpolls}

\pagebreak

\section{Motivation}

As discussed in~\cite{P3948R0}, because of language inconsistencies, \std\code{constant_wrapper}
is inconsistently not unwrapping for call and subscript operators whereas all other operators can
be found via ADL and the conversion operator implemented in \code{constant_wrapper}.
Looking at \std\code{reference_wrapper} we see that this same issue has been resolved via an
\code{operator()} overload that unwraps and applies \code{INVOKE}.

I always had these unwrapping overloads in my implementation of \code{constant_wrapper} shipping in
the vir-simd library\footnote{https://github.com/mattkretz/vir-simd}
(\code{vir::constexpr_wrapper}).
While replacing my implementation with \std\code{constant_wrapper} I noticed the mismatch.

\subsection{Design Principle}

The following has always been my thinking on what \code{constant_wrapper} is / needs to be, which
informed all my opinions on its API\@:
\code{constant_wrapper} exists to be able to \emph{use function arguments in place of template
arguments}.
\begin{itemize}
  \item Passing a \code{1} as a template argument, the function sees an \code{int}.
  \item Passing a \code{cw<1>} as a function argument, the function does not see an \code{int} but a
    \code{constant_wrapper<1, int>}; I actually wanted an \code{int}.
\end{itemize}
Therefore, \code{constant_wrapper<X>} should transparently unwrap into \code{X} whenever possible
(similar to \code{reference_wrapper} unwrapping to the reference it holds), except if it \emph{can}
stay in the type space, in which case it unwraps all operands/arguments and re-wraps the result.

\begin{center}
\fcolorbox{RoyalBlue}{White!90!RoyalBlue}{\larger%
it's the thing it holds, unless it can stay wrapped
}
\end{center}

This leads to the following expectations:
\smallskip\begin{lstlisting}
constexpr int iota[4] = {0, 1, 2, 3};
\end{lstlisting}
\begin{tabular}{rcll}
  \code{cw<1> + 1} & $\rightarrow$ & \code{2} & the thing it holds \\
  \code{cw<1> + cw<1>} & $\rightarrow$ & \code{cw<2>} & can stay wrapped \\
  \code{cw<iota>[1]} & $\rightarrow$ & \code{1} & the thing it holds \\
  \code{cw<iota>[cw<1>]} & $\rightarrow$ & \code{cw<1>} & can stay wrapped \\
  \code{cw<fun>(1, 2)} & $\rightarrow$ & \code{fun(1, 2)} & the thing it holds \\
  \code{cw<fun>(cw<1>, cw<2>)} & $\rightarrow$ & \code{cw<fun(1, 2)>} & can stay wrapped$^*$ \\
  \code{cw<fun>(cw<1>, 2)} & $\rightarrow$ & \code{fun(cw<1>, 2)} & the thing it holds \\
  \code{cw<unary>()} & $\rightarrow$ & \code{cw<unary()>} & can stay wrapped$^*$
\end{tabular}

\medskip
$^*$ \code{fun(1, 2)} and \code{unary()} need to be constant expressions, otherwise unwrap.

%\vspace{0pt plus 2em}
\subsection{Status Quo Examples}
\medskip
\begin{lstlisting}
auto test1()
{
  constexpr int iota[4] = {0, 1, 2, 3};
  auto x = std::cw<iota>;
  return x[1]; // #1 OK
}

auto test2()
{
  auto x = std::cw<std::array<int, 4> {0, 1, 2, 3}>;
  return x[1]; // #2 ill-formed
}
\end{lstlisting}

The subscript in \code{\#1} is fine because \code{x} is convertible to \code{int[4]} and because
operator lookup is different for built-in types, the built-in subscript operator is found and
\code{int(1)} is returned.
The subscript in \code{\#2} does not work because, even though \code{array<int, 4>} is an associated
namespace and \code{x} is convertible to \code{array<int, 4>}, the \code{array::operator[]} member
function is not found.
This is because only hidden friend operators are considered via ADL\@.
Note how ADL makes the following work for \code{operator+}:
\medskip
\begin{lstlisting}
struct X { friend int operator+(X, int a) { return a + 1; } };

auto test3()
{
  auto x = std::cw<X{}>;
  return x + 1; // #3 OK
}
\end{lstlisting}
The expression \code{x + 1} in line \code{\#3} finds \code{X::operator+} via ADL and thus converts
the left operand to \code{operator+} to \code{X}, returning \code{int(2)}.

The situation for \code{operator()} is equivalent:
\medskip
\begin{lstlisting}
int fun(int x) { return x + 1; }

auto test1()
{
  auto x = std::cw<fun>;
  return x(1); // #4 OK
}

auto test2()
{
  auto x = std::cw<[](int x) { return x + 1; }>;
  return x(1); // #5 ill-formed
}
\end{lstlisting}
The call expression in \code{\#4} relies on special core wording that looks through the conversion
operator to find the function pointer and then call the function.
The expression in \code{\#5}, however, cannot find the \code{operator()} member of the lambda,
because the operator is a member, not a hidden friend.

\section{Discussion}

The most glaring question on this issue is why would we do this for \code{operator()} and
\code{operator[]} but for none of the other operators.
This seems inconsistent.
However, we need to realize that the inconsistency is in the language, forcing the inconsistent
definition of operators in the library.
This makes the \emph{behavior} of the API, where operators unwrap the value of the
\code{constant_wrapper} transparently, more consistent.
After all, the conversion operator and the additional type template argument in
\code{constant_wrapper} exist exactly because \code{constant_wrapper} is supposed to transparently
unwrap.

Wouldn't it then be better to fix the language?\\
For what it's worth, I think it would be a hugely helpful change to make the behavior of all
operators as consistent as possible.
Currently, every operator has its own set of restrictions and extras.
If I could, I'd make \emph{every} operator overloadable as non-member (and thus hidden friend) and
implement all operators of standard library types as hidden friends.
But the amount of code we would break \ldots{}
The only possibility that is not a breaking change is to add syntax that opts into new behavior,
which has a high acceptance barrier.

In terms of consistency we also have \std\code{reference_wrapper} to consider.
The similar naming is not accidental.
Consequently, the unwrapping behavior should also be consistent.

\subsection{Exploration of potential ambiguities}

Consider the examples discussed in \textcite{P3792R0}, which overload on \code{constant_wrapper} and
the type it is convertible to.
We can construct a similar case with the existing \code{constant_wrapper} operator overloads:
\medskip\begin{lstlisting}
struct V {
  int data;

  constexpr V(int x) : data(x) {}

  template <auto N>
  constexpr V(std::constant_wrapper<N, int> x) : data(x + 1) {}

  friend constexpr V operator+(const V& a, const V& b) {
    return a.data + b.data;
  }
};

auto a = cw<V(1)> + 2;            //    V(3)
auto b = cw<V(1)> + cw<2>;        // cw<V(3)>
auto c = cw<V(1)> + cw<V(2)>;     // cw<V(3)>
auto d = cw<V(1)> + cw<cw<2>>;    // cw<V(4)>
auto e = cw<V(1)> + cw<cw<V(2)>>; // cw<V(3)>
\end{lstlisting}

\code{V} is structural, convertible from \code{int} and from \code{cw<N>} (where \code{N} is of type
\code{int}).
\begin{enumerate}
  \item[\code{a}.] ADL finds \code{V::operator+} and \code{constant_wrapper::operator+}, but only
    the former is viable.
    The result is \code{V(3)} (“it's the thing it holds”).

  \item[\code{b}.] ADL again finds both operators, but now \code{constant_wrapper::operator+} is
    viable and requires no conversions.
    The \code{constant_wrapper} operator evaluates \code{V(1) + 2}, which is \code{V(3)}.
    The result is thus \code{cw<V(3)>} (“it can stay wrapped”).

  \item[\code{c}.] See \code{b}.

  \item[\code{d}.] ADL still finds both operators, but now \code{V::operator+} isn't even viable
    anymore.
    \code{constant_wrapper::operator+} evaluates \code{V(1) + cw<2>}, which calls
    \code{V::operator+} and requires a conversion from \code{cw<2>} to \code{V}.
    The conversion constructor adds \code{1} and \code{V(1) + cw<2>} thus equals \code{V(4)}.
    It is then wrapped into \code{constant_wrapper} again, making the result \code{cw<V(4)>} (“it
    can stay wrapped” and the inner evaluation is “the thing it holds”).

  \item[\code{e}.] See \code{d}, except that we don't need a conversion from \code{cw<2>} to
    \code{V} and thus the result is \code{cw<V(3)>}.
\end{enumerate}

We observe that the \code{constant_wrapper} “it can stay wrapped” operator takes precedence over the
\code{V} “it's the thing it holds” operator.
The same behavior is now proposed for call and subscript.
Consider:

\medskip\begin{lstlisting}
struct Callable
{
  static constexpr int operator()(int) {
    return 1;
  }

  template <auto N>
  static constexpr int operator()(std::constant_wrapper<N, int>) {
    return 2;
  }
};

auto a = Callable{}(0);             // 1
auto b = Callable{}(cw<0>);         // 2
auto c = cw<Callable{}>(0);         // 1
auto d = cw<Callable{}>(cw<0>);     // cw<1>
auto e = cw<Callable{}>(cw<cw<0>>); // cw<2>
\end{lstlisting}

\code{a} and \code{b} are obvious.
\code{c} is new with this paper and also fairly obvious (“it's the thing it holds”).
\code{d} is the case one might consider ambiguous.
Do we expect \code{cw<Callable{}>} to unwrap and then call the \code{constant_wrapper} overload?
Or do we expect both to unwrap, call the \code{int} overload and wrap it again?
Above, we observed that “it can stay wrapped” takes precedence over “it's the thing it holds”.
\code{d} being deduced as \code{cw<1>} is consistent and follows this one simple rule.
\code{e}, finally uses the \code{constant_wrapper} call operator to call \code{Callable{}(cw<0>)},
which is \code{2}, wraps it again and thus is deduced as \code{cw<2>}.

If we construct a 2-argument call operator, then all operands need to be \code{constant_wrapper}
objects for “it can stay wrapped” to take precedence.
A call like \code{cw<Callable2arg{}>(cw<1>, 2)} cannot produce a \code{constant_wrapper} (because
\code{2} is a function argument and therefore not a constant expression in the operator body) and
thus the expression evaluates as \code{Callable2arg(cw<1>, 2)}.

\subsection{Example: Interaction with \code{bind_front} / \code{bind_back}}

With this paper \std\code{bind_front(\std{}cw<\&T::memfun>, obj)} becomes useful.
Without it, \std\code{bind_front(\std{}cw<\&T::memfun>, obj)} itself is well-formed, but calling
\code{operator()} on it isn't.

Consider:\medskip
\begin{lstlisting}
struct X {
  int fun(int x);
};

int blackhole = 0;

void eval(auto&& b) { blackhole = b(1); }

void test1(X& t) {
  eval(std::bind_front(&X::fun, t)); // indirect call in eval
}

void test2(X& t) {
  eval(std::bind_front<&X::fun>(t)); // direct call in eval
}

void test3(X& t) {
  eval(std::bind_front(std::cw<&X::fun>, t)); // direct call in eval
}
\end{lstlisting}

All three are equivalent.
However, \code{test3} compiles to the same code as \code{test2}\footnote{tested on GCC trunk with
libstdc++} while \code{test1} is less efficient since it uses an indirect call in \code{eval}.

Finally, let's consider a realistic example where \code{bind_back<fn>(...)} vs.
\code{bind_back(cw<fn>, ...)} makes an observable difference.
We use \std\code{bind_back} to create a unary transformation that returns a fifth for a given input.
The \code{Scaler} implementation uses a \code{constant_wrapper} overload to optimize division for
statically known divisors.

\medskip\begin{lstlisting}
template<typename T>
  struct Scaler {
    static constexpr T operator()(T v, T d) // #1
    { return v / d; }

    // optimization of the above: division by known constant doesn't need
    // slow DIV instruction
    template<auto D>
    static constexpr T operator()(T v, std::constant_wrapper<D, T> d) // #2
    { return v / d; }
  };

template<template<typename ...> class Tuple, typename... Args, typename S>
constexpr auto transform_tuple(Tuple<Args...> t, S&& s) {
  auto [... elems] = std::move(t);
  return Tuple{ s(elems)... }; // CTAD
}

constexpr std::strided_slice slice {15, std::cw<20>, std::cw<5>};
auto sl1 = transform_tuple(slice, std::bind_back<Scaler<int>{}>(std::cw<5>));
auto sl2 = transform_tuple(slice, std::bind_back(std::cw<Scaler<int>{}>, std::cw<5>));
\end{lstlisting}

\begin{itemize}
  \item [\code{sl1}:] \lstinline@std::strided_slice {3, 4, 1}@\\
    The calls to \code{s(elems)...} in \code{transform_tuple} all invoke overload \code{\#2}.
    For \code{s(elems...[0])} overload \code{\#2} is a perfect match.
    \code{s(elems...[1])} calls \code{Scaler<int>{}(cw<20>, cw<5>)} for which overload \code{\#2} is
    a better match (only one conversion).
    Same for \code{s(elems...[2])}.

  \item [\code{sl2}:] \lstinline@std::strided_slice {3, std::cw<4>, std::cw<1>}@\\
    \code{s(elems...[0])} in \code{transform_tuple} calls \code{cw<Scaler<int>\{\}>(15, cw<5>)}
    which invokes overload \code{\#2}.
    \code{s(elems...[1])} calls \code{cw<Scaler<int>\{\}>(cw<20>, cw<5>)}.
    Here all operands to the expression are \code{constant_wrapper}s and \code{operator() \#1} is
    declared \code{constexpr}, which allows the “it can stay wrapped” behavior.
    The result of \code{s(elems...[1])} therefore is \code{cw<4>}.
    Likewise, \code{s(elems...[2])} evaluates to \code{cw<1>}.
\end{itemize}

In this example, a change from \code{bind_back<Callable>(cw<5>)} to \code{bind_back(cw<Callable>,
cw<5>)} leads to a \emph{different} result.
This is not necessarily wrong or bad.
It can certainly be argued that the \code{sl2} result is better.
It can also be argued that the \code{sl2} result is surprising.
However, it still follows the simple rule of “it's the thing it holds, unless it can stay wrapped”.

\subsection{\placeholder{INVOKE} should be used (or not used) consistently}

\cite{P3978R1} proposed to use \placeholder{INVOKE} for the unwrapping call operator and otherwise
apply \code{operator()} directly on \code{value}.
This is inconsistent and not proposed anymore.
Instead this paper proposes to use \placeholder{INVOKE} in both cases.
Relative to \cite{P3978R1}, this paper's wording has the following change:
\begin{wgText}\setcounter{Paras}{4}
\begin{itemdescr}
\pnum
  Let \placeholder{call-expr} be
  \tcode{constant_wrapper<\wgChange{value(}{\placeholder{INVOKE}(value, }remove_reference_t<Args>::value...)>\{\}} if all types in
  \tcode{remove_reference_t<Args>...} satisfy \exposconcept{constexpr-param} and
  \tcode{constant_wrapper<\wgChange{value(}{\placeholder{INVOKE}(value, }remove_reference_t<Args>::value...)>} is a valid type, otherwise
  let \placeholder{call-expr} be \tcode{\placeholder{INVOKE}(value, std::forward<Args>(args)...)}.
\end{itemdescr}
\end{wgText}
Example:
\medskip\begin{lstlisting}[style=Vc]
struct T {
  int v;
  constexpr int add(int x) { return v + x; }
};

auto a = cw<&T::add>(cw<T{1}>, cw<2>) // -> cw<3>
\end{lstlisting}

This seems to be the right choice for consistency.
However, we did not use \placeholder{INVOKE} initially for \code{constant_wrapper} because the
intent was to model the exact interface of the underlying type.
Using \placeholder{INVOKE}, we are deviating from that choice, giving the call operator more
functionality than the type it wraps.
This is also true for the “it's the thing it holds” case.
The other choice for resolving the inconsistency is the removal of \placeholder{INVOKE} from the
\code{constant_wrapper} wording altogether:
\begin{wgText}\setcounter{Paras}{4}
\begin{itemdescr}
\pnum
  Let \placeholder{call-expr} be
  \tcode{constant_wrapper<\wgChange{\placeholder{INVOKE}(value, }{value(}remove_reference_t<Args>::value...)>\{\}} if all types in
  \tcode{remove_reference_t<Args>...} satisfy \exposconcept{constexpr-param} and
  \tcode{constant_wrapper<\wgChange{\placeholder{INVOKE}(value, }{value(}remove_reference_t<Args>::value...)>} is a valid type, otherwise
  let \placeholder{call-expr} be \tcode{\wgChange{\placeholder{INVOKE}(value, }{value(}std::forward<Args>(args)...)}.
\end{itemdescr}
\end{wgText}

\section{Header for \code{constant_wrapper}}

We should reconsider the placement of \code{constant_wrapper} in \code{<type_traits>}.
\begin{enumerate}
  \item With this change we use \emph{INVOKE}, which typically isn't available already in
    \code{<type_traits>}.

  \item \code{constant_wrapper} is not a trait.
    \code{integral_constant} is not a trait either, but it is the type of many traits.
    \code{constant_wrapper} is a utility (\code{<utility>}).

  \item \code{<functional>}? Too heavy, in my opinion.
\end{enumerate}

For reference in libstdc++ we have the following dependencies:
\medskip\begin{lstlisting}
> grep include type_traits
# include <bits/c++0x_warning.h>
#include <bits/c++config.h>

> grep include utility
#include <bits/c++config.h>
#include <bits/stl_relops.h>
#include <bits/stl_pair.h>
#include <type_traits>
#include <bits/move.h>
#include <initializer_list>
#include <ext/numeric_traits.h>

> grep include functional
#include <bits/c++config.h>
#include <bits/stl_function.h>
#include <new>
#include <tuple>
#include <type_traits>
#include <bits/functional_hash.h>
#include <bits/invoke.h>
#include <bits/refwrap.h>       // std::reference_wrapper and _Mem_fn_traits
#include <bits/std_function.h>  // std::function
# include <unordered_map>
# include <vector>
# include <array>
# include <utility>
# include <bits/stl_algo.h>
# include <bits/ranges_cmp.h>
# include <compare>
\end{lstlisting}

\section{Future Work?}

With this paper, \code{constant_wrapper} almost achieves parity with what was proposed for
\code{fn_t} / \code{function_wrapper} by \textcite{P3774R0} and \textcite{P3843R2}.
The one feature still missing is the conversion to a different function pointer type via implicit
creation of a thunk.
This could be added as a member function (\code{to_func()}), which returns a different type that
implements the conversion operator as shown in Section 4.2 of \cite{P3774R0}.


\section{Wording}
\subsection{Feature test macro}

In [version.syn] bump the \code{__cpp_lib_constant_wrapper} version.

\subsection{Modify [const.wrap.class]}

In [const.wrap.class], change:

\begin{wgText}[{[const.wrap.class]}]
\begin{codeblock}
    template<@\exposconcept{constexpr-param}@ L, @\exposconcept{constexpr-param}@ R>
      friend constexpr auto operator->*(L, R) noexcept -> constant_wrapper<L::value->*(R::value)>
        { return {}; }

    @\wgRem{\rmfamily\itshape// call and index}@
    @\wgRem{template<\exposconcept{constexpr-param} T, \exposconcept{constexpr-param}... Args>}@
      @\wgRem{constexpr auto operator()(this T, Args...) noexcept}@
        @\wgRem{requires requires \{ constant_wrapper<T::value(Args::value...)>(); \}}@
          @\wgRem{\{ return constant_wrapper<T::value(Args::value...)>\{\}; \}}@
    @\wgRem{template<\exposconcept{constexpr-param} T, \exposconcept{constexpr-param}... Args>}@
      @\wgRem{constexpr auto operator[](this T, Args...) noexcept}@
        @\wgRem{-> constant_wrapper<(T::value[Args::value...])>}@
          @\wgRem{\{ return \{\}; \}}@

    // pseudo-mutators
    template<@\exposconcept{constexpr-param}@ T>
      constexpr auto operator++(this T) noexcept
        -> constant_wrapper<++Y> { return {}; }
\end{codeblock}

[\dots]

\begin{codeblock}
  };

  template<@\exposid{cw-fixed-value}@ X, class>
  struct constant_wrapper : @\exposid{cw-operators}@ {
    static constexpr const auto & value = X.@\exposid{data}@;
    using type = constant_wrapper;
    using value_type = decltype(X)::@\exposid{type}@;

    template<@\exposconcept{constexpr-param}@ R>
      constexpr auto operator=(R) const noexcept
        -> constant_wrapper<X = R::value> { return {}; }

    constexpr operator decltype(auto)() const noexcept { return value; }

    @\wgAdd{template<class... Args>}@
      @\wgAdd{static constexpr decltype(auto) operator()(Args\&\&... args) noexcept(\seebelow)}@
    @\wgAdd{template<class... Args>}@
      @\wgAdd{static constexpr decltype(auto) operator[](Args\&\&... args) noexcept(\seebelow)}@
  };
}
\end{codeblock}

[\ldots]

\begin{itemdecl}
constexpr @\exposid{cw-fixed-value}@(T (&arr)[Extent]) noexcept;
\end{itemdecl}

\setcounter{Paras}{3}

\begin{itemdescr}
\pnum
\effects
Initialize elements of \exposid{data} with corresponding elements of \tcode{arr}.
\end{itemdescr}

\begin{itemdecl}
@\wgAdd{template<class... Args>}@
  @\wgAdd{static constexpr decltype(auto) operator()(Args\&\&... args) noexcept(\seebelow)}@
\end{itemdecl}

\begin{itemdescr}
\pnumAdd
\wgAdd{%
  Let \placeholder{call-expr} be
  \tcode{constant_wrapper<\placeholder{INVOKE}(value, remove_reference_t<Args>::value...)>\{\}} if all types in
  \tcode{remove_reference_t<Args>...} satisfy \exposconcept{constexpr-param} and
  \tcode{constant_wrapper<\placeholder{INVOKE}(value, remove_reference_t<Args>::value...)>} is a valid type, otherwise
  let \placeholder{call-expr} be \tcode{\mbox{\placeholder{INVOKE}}(value, std::forward<Args>(args)...)}.
}

\pnumAdd
\wgAdd{%
\constraints
\placeholder{call-expr} is a valid expression.
}

\pnumAdd
\wgAdd{%
\effects
Equivalent to: \tcode{return \placeholder{call-expr};}
}

\pnumAdd
\wgAdd{%
\remarks
The expression inside \tcode{noexcept} is \tcode{noexcept(\placeholder{call-expr})}.
}
\end{itemdescr}

\begin{itemdecl}
@\wgAdd{template<class... Args>}@
  @\wgAdd{static constexpr decltype(auto) operator[](Args\&\&... args) noexcept(\seebelow)}@
\end{itemdecl}

\begin{itemdescr}
\pnumAdd
\wgAdd{%
  Let \placeholder{subscr-expr} be
  \tcode{constant_wrapper<value[remove_reference_t<Args>::value...]>\{\}} if all types in
  \tcode{remove_reference_t<Args>...} satisfy \exposconcept{constexpr-param} and
  \tcode{constant_wrapper<value[remove_reference_t<Args>::value...]>} is a valid type, otherwise
  let \placeholder{subscr-expr} be \tcode{value[std::forward<Args>(args)...]}.
}

\pnumAdd
\wgAdd{%
\constraints
\placeholder{subscr-expr} is a valid expression.
}

\pnumAdd
\wgAdd{%
\effects
Equivalent to: \code{return \placeholder{subscr-expr};}
}

\pnumAdd
\wgAdd{%
\remarks
The expression inside \tcode{noexcept} is \tcode{noexcept(\placeholder{subscr-expr})}.
}
\end{itemdescr}

\end{wgText}

\end{document}
% vim: sw=2 sts=2 ai et tw=100
