\section{Changelog}
\begin{revision}
\item Constrain the \code{consteval} ctor to arithmetic types that naturally
  convert to the \code{basic_vec}'s value-type (checked via \code{common_type}).
\item Discuss consequence on \code{common_type} (besides \code{convertible_to}).
\item Discuss consequence on [simd.math].
\item Discuss \code{consteval} ctor as immediate-escalating expression.
\item Discuss potential consequences for [simd.math].
\end{revision}

\begin{revision}
\item Remove proposed polls.
\item Update wording to remove unnecessary \exposid{simd-broadcast-arg}.
\item Add [simd.math] wording changes to match original design intent (as if it
  were an explicit overload set).
\item Drive-by wording fixes/improvements to [simd.math]:
  \begin{itemize}
    \item \exposid{math-common-simd-t<V0, V1>} did not work for non-default ABI
      tag combinations with scalars.
    \item Spell out precondition on integral \code{abs}.
  \end{itemize}
\end{revision}

\begin{revision}
\item Restore incorrectly removed \exposid{math-floating-point} wording.
\item Fix incorrect constraint in \code{hypot} expample.
\item Split wording section into the \code{consteval} ctor part and the [simd.math] part.
\end{revision}

\begin{revision}
\item Move \code{consteval} constructor discussion and wording into a separate paper \parencite{P4012R0}.
\item In [simd.math] wording, move all functions with the same name together.
\item Fix return type of additional overloads of \code{isgreater},
  \code{islessequal}, \code{islessgreater}, and \code{isunordered}.
\item Add missing overloads for \code{isgreaterequal}, \code{isless}.
\item Simplify placeholder functions to \code{static_cast} to prvalue rather than const-ref.
  Since \code{basic_vec} types are trivially copyable a copy can be elided by
  the implementation as-if.
\end{revision}
