\section{Changelog}
\begin{revision}
  \item Added \code{bounded::integer} motivation and example.
  \item Added a reference to P0927R0; making a stronger case for the simple choice.
\end{revision}

\begin{revision}
  \item Discuss \code{common_type}.
  \item Discuss overloading \code{operator?:(bool, ...)}.
  \item Mention \type{chrono::duration} and other numeric types as motivation.
\end{revision}

\begin{revision}
  \item Add Tony tables.
  \item Explore defaulting \code{operator?:}.
  \item Discuss synthesizing \code{operator?:} from \code{common_type}.
  \item Define a \code{contextual_boolean} concept, that most overloads should use instead of a naïve \bool parameter.
  \item Try to be clearer about generality, consistency, and orthogonality of this proposal.
  \item Add \code{boost::yap} as another existing library that is missing the ability to overload \code{?:}.
  %\todo Add a tony table for the design choice “specialize \code{common_type}” vs. “define conditional operator overload”.
  %\todo Discuss GCC implementation.
\end{revision}

\begin{revision}
  \todo Explore: \code{operator?:} for contextual-bool automatically uses \code{common_type} and cannot be overloaded otherwise.
        It does deferred evaluation.
        Non-boolean conditions (i.e. the type of the first argument is not contextually convertible to \bool) can be overloaded and are therefore a normal function call without deferred evaluation.
  \todo \code{auto operator?:(T, U) = default} could mean \code{common_type_t<T, U> operator?:(...)}, i.e. synthesize a builtin operator using \code{common_tpye} for the return type.
  \todo Explore: Definition of \code{operator?:(Cond, A, B)} automatically synthesizes \code{operator?:(Cond, B, A)} calling the former.
        (Note, there are no compatibility issues like for defaulted compares.)
  \todo More motivation: P2655R1 "\code{common_reference_t} of \code{reference_wrapper} should be a reference type".
        P2655R1 resolves the problem via another template specialization. Instead making \code{operator?:} do the right thing would be more generic and an obvious fix for the type's semantics.
        \todo Units example:
        \begin{lstlisting}[style=Vc]
        quantity<isq::speed[km / h]> speed_1 = 42 * isq::speed[m / s];
        quantity<isq::speed[m / s]> speed_2 = 42 * isq::speed[km / h];
        quantity_of<isq::speed> auto speed_3 = speed_1 + speed_2;

        using ct = std::common_type_t<decltype(speed_1), decltype(speed_2)>;

        static_assert(std::is_same_v<ct, decltype(speed_3)>);
        static_assert(std::is_same_v<ct, quantity<reference<
          isq::speed, scaled_unit<mag<ratio{1, 18}>, derived_unit<si::metre,
                                                       per<si::second>>>{}>{}>>);

        // this cannot but should work:
        quantity_of<isq::speed> auto = cond ? speed_1 : speed_2;
        \end{lstlisting}
\end{revision}

  %\todo Discuss implications for “contextually convertible to \bool” implementations.
