\newcommand\wgTitle{Exploration of namespaces for std::simd}
\newcommand\wgName{Matthias Kretz <m.kretz@gsi.de>}
\newcommand\wgDocumentNumber{D3287R3}
\newcommand\wgGroup{LWG}
\newcommand\wgTarget{\CC{}26}
\newcommand\wgAcknowledgements{Daniel Towner and Ruslan Arutyunyan contributed to this paper via discussions / reviews.}

\usepackage{mymacros}
\usepackage{wg21}
\setcounter{tocdepth}{2} % show sections and subsections in TOC
\hypersetup{bookmarksdepth=5}
\usepackage{changelog}
\usepackage{underscore}
\usepackage{multirow}

\addbibresource{extra.bib}

\newcommand\simd[1][]{\type{ba\-sic\_simd#1}\xspace}
\newcommand\simdT{\type{ba\-sic\_simd\MayBreak<\MayBreak{}T>}\xspace}
\newcommand\valuetype{\type{val\-ue\_type}\xspace}
\newcommand\referencetype{\type{ref\-er\-ence}\xspace}
\newcommand\mask[1][]{\type{ba\-sic\_simd\_mask#1}\xspace}
\newcommand\maskT{\type{ba\-sic\_simd\_mask\MayBreak<\MayBreak{}T>}\xspace}
\newcommand\wglink[1]{\href{https://wg21.link/#1}{#1}}
\newcommand\simdgeneric{SIMD-gener\-ic\xspace}

\newcommand\nativeabi{\UNSP{native-abi}}
\newcommand\deducet{\UNSP{deduce-t}}
\newcommand\simdsizev{\UNSP{simd-size-v}}
\newcommand\simdsizetype{\UNSP{simd-size-type}}
\newcommand\simdselect{\UNSP{simd-select-impl}}
\newcommand\maskelementsize{\UNSP{mask-element-size}}
\newcommand\integerfrom{\UNSP{integer-from}}
\newcommand\constexprwrapperlike{\UNSP{constexpr-wrapper-like}}
\newcommand\vectorizable{\UNSP{vectorizable}}

\renewcommand{\lst}[1]{Listing~\ref{#1}}
\renewcommand{\sect}[1]{Section~\ref{#1}}
\renewcommand{\ttref}[1]{Tony~Table~\ref{#1}}
\renewcommand{\tabref}[1]{Table~\ref{#1}}

\newcommand\myrating[1]{\par\noindent{\color{Headings}\scshape{}my rating:} #1}

\begin{document}
\selectlanguage{american}
\begin{wgTitlepage}
  In recent discussions about \code{simd} in LEWG, notably on 2023-06-16 while
  discussing \code{permute}, \code{expand}, and \code{compress}, there was a
  request for a paper exploring placing all \code{simd} non-member functions
  into a sub-namespace.
  \ldots or potentially any other means of using namespaces to improve the
  \code{simd} API.

  This paper explores a few ideas.
\end{wgTitlepage}

\pagestyle{scrheadings}

\input{changelog}
\input{strawpolls}

\section{Introduction \& Motivation}

Using the example of \std\code{permute(\simd, idx_perm)}, one of the unavoidable
LEWG discussions/decisions is about whether \code{simd} can grab the name
“permute”, potentially blocking its use for other facilities in the standard
library.\footnote{Just to clarify, I agree with the concern and I feel uneasy
with the need for \code{simd} to grab as many names as it would need to.}
With P3067R0 (“Provide named permutation functions for \std\code{simd}”), the
list of non-member functions to add to \std\ becomes:
\code{permute}, \code{expand}, \code{compress}, \code{grow}, \code{stride},
\code{chunk}, \code{reverse}, \code{re\-peat_all}, \code{re\-peat_each},
\code{transpose}, \code{zip}, \code{unzip}, \code{cat}, \code{ex\-tract},
\code{rotate}, \code{shift_left}, \code{shift_right}, and \code{align}.
All of these names would likely need a \code{simd} prefix if they want to go
into \std.

And then we're adding \simd overloads for all of \code{<cmath>} and
\code{<bit>}, \ldots.

So we need to understand whether there are viable alternatives to \code{simd}
naming.
This paper tries to explore the field as far as I believe is still sensible.
The goal is to come up with a consistent naming strategy for everything related
to \code{simd}.

\subsection{Motivation for a \code{std::simd} namespace}
\label{sec:namespacemotivation}

A \std\code{simd} namespace encapsulates everything related to data-parallel
types, creating an easy-to-explore, isolated space for these functionalities.
It avoids the need for (inconsistent) \code{simd_} prefixes by grouping related
functions into their own namespace.
It increases flexibility for future extensions to be organized within the
\std\code{simd} namespace.
Users can easily alias the namespace (e.g. \code{namespace simd = std::simd}),
reducing verbosity in the code but maintaining a clean and logical structure.

The \std\code{simd} namespace approach might be cleaner in the long term.
It avoids overloads like \code{std::reduce} and \code{std::all_of}, reducing
the likelihood of confusion.
While a \code{simd::simd} duplication in naming would be a bit awkward, it is a
one-time issue that users can learn to live with.
And we should consider renaming the alias template to
\std\code{simd::\MayBreak{}vec} (\std\code{simd::\MayBreak{}basic_vec} for the
class template).

This approach also aligns with other \CC{} standard library features,
where sub-namespaces are used to logically group related functionality:\\
\std\code{chrono},
\std\code{execution},
\std\code{filesystem},
\std\code{linalg},
\std\code{numbers},
\std\code{pmr},\\
\std\code{ranges}, and
\std\code{ranges::views}/\std\code{views}.

\subsection{\simdgeneric programming}

In this paper I want to use the term \emph{\simdgeneric} programming.
Note that in the space of types, \simdT is a generalization of \code{T} or ---
vice-versa --- \code{T} is the degenerate case of \simdT.
(The same is true for \mask and \code{bool}.)
We've touched upon this when we talked about regularity and how \simdT is
designed to retain regularity of each individual element inside the \simd,
leading to something I called “data-parallel regularity” of \simdT, for lack of
an existing term.

The \code{simd} design aims to allow users to replace \code{T} with \simdT in
their code without requiring any further code changes.
If this works (and because of branching on individual values of \code{T} it
cannot work for all code) I call such code \simdgeneric.

The following text uses this term because the use of namespaces opens an
interesting facility to opt in and out of some aspects of \simdgeneric
programming.

\section{Exploration}

When exploring naming and namespacing, I use the following functions to showcase the effect.
I then try to come up with all ways to use and abuse the facilities.
In addition I mention the effect of the choice on \simdgeneric programming.
To complete the picture, I added a concept that seems like something we might
want to add in the future, but for which there is currently no proposal coming
forward.

Note: we have to discuss the range vs. iterator argument to load/gather seperately.
This paper does not explore the issue.
I also removed \code{constexpr} and \code{noexcept} since they are irrelevant
to the exploration at hand.

\begin{enumerate}
  \item \label{fun:generator} \simd generator

    Status quo (P1928R9):
    \medskip\begin{lstlisting}[style=Vc]
std::simd<int> iota([](int i) { return i; });
    \end{lstlisting}

  \item \label{fun:load} \simd load from contiguous range

    Status quo (P1928R9):
    \medskip\begin{lstlisting}[style=Vc]
std::vector<int> data = {...};
std::simd<int> chunk(data.begin());
    \end{lstlisting}

  \item \label{fun:gather} \simd gather from contiguous range

    Status quo (P2664R6):
    \medskip\begin{lstlisting}[style=Vc]
std::vector<int> data = {/*...*/};
std::simd<int> idxs = /*...*/;
std::simd<int> std::gather_from(data, idxs);
    \end{lstlisting}

  \item \label{fun:permute} \simd permutations

    Status quo (P2664R6):
    \medskip\begin{lstlisting}[style=Vc]
std::simd<int> v = /*...*/;
std::simd<int> v2 = std::permute(v, [](int i) { return i ^ 1; });
    \end{lstlisting}

  \item \label{fun:select} \simd ternary operator replacement

    Status quo (P1928R9):
    \medskip\begin{lstlisting}[style=Vc]
std::simd<int> v = /*...*/;
std::simd<int> abs = std::simd_select(v >= 0, v, -v);
    \end{lstlisting}

  \item \label{fun:math} Math functions and algorithms

    Status quo (P1928R9):
    \medskip\begin{lstlisting}[style=Vc]
std::simd<float> x = /*...*/;
std::simd<float> y = std::exp(x);
std::simd<float> z = std::min(x, y);
    \end{lstlisting}

  \item \label{fun:maskred} Mask reductions

    Status quo (P1928R9):
    \medskip\begin{lstlisting}[style=Vc]
std::simd<float> x = /*...*/;
if (std::all_of(x > 0)) /*...*/
    \end{lstlisting}

  \item \label{concept} Simd concepts
    \begin{itemize}
      \item Constrain whether a type is a \simdT with \std\code{integral<T>}.
      \item Constrain whether a type is either \std\code{integral} or a \simdT
        with \std\code{integral<T>}.
    \end{itemize}

\end{enumerate}

\subsection{Status quo (latest revision of \code{simd} papers)}
\begin{description}
  \item[pros]
    \begin{itemize}
      \item \std\code{simd} is as concise as it could possibly be.

      \item Fairly good support for \simdgeneric programming.
    \end{itemize}

  \item[cons]
    \begin{itemize}
      \item We have a mix of non-member functions with and without
        \code{simd_} prefix.

      \item Most non-member functions would be nicer to read in code without
        the \code{simd} prefix.
        We introduce the prefix only because we are wary of the “name grab”
        in \code{std}.
        \Ie the motivation for the current naming scheme isn't the design of
        the \code{simd} API, but the freedom to evolve the standard library
        in the future.

      \item Load and gather (which are very similar in loading a SIMD
        “register” from a contiguous range of values) are inconsistent: One
        uses a constructor and member function, the other only a non-member
        function.

      \item Loads, stores, and the \code{simd} generator constructor cannot
        be used in \simdgeneric code.
    \end{itemize}
\end{description}

\subsection{Explorations in previous revision(s) of this paper}
\begin{itemize}
  \item Alternative 1: every function is a non-member with \code{simd} prefix
  \item Alternative 2: every function is a non-member without \code{simd} prefix
  \item Alternative 3: place everything but types into a namespace
  \item Alternative 4: make all non-member functions hidden friends
\end{itemize}

\input{alt5}
\pagebreak\input{alt6}
\pagebreak\input{alt7}
\pagebreak\input{alt8}

\input{naming}

\pagebreak
\section{Recommendation: Two examples after renaming}

Option 1:
\medskip\begin{lstlisting}
namespace dp = std::datapar;

// compute log for positive inputs
dp::simd<float> f(std::span<float> data)
{
  dp::simd<float>  x = dp::load_from(data);
  dp::mask<float> positive = x > 0.f;
  dp::simd<float>  l = std::log(dp::select(positive, x, 1.f));
  return dp::select(positive, l, x);
}
\end{lstlisting}

Option 2:
\medskip\begin{lstlisting}
namespace simd = std::simd;

// compute log for positive inputs
simd::vec<float> f(std::span<float> data)
{
  simd::vec<float>  x = simd::load_from(data);
  simd::mask<float> positive = x > 0.f;
  simd::vec<float>  l = std::log(simd::select(positive, x, 1.f));
  return simd::select(positive, l, x);
}
\end{lstlisting}

Compare against the status quo:
\medskip\begin{lstlisting}
// compute log for positive inputs
std::simd<float> f(std::span<float> data)
{
  std::simd<float>  x = std::simd_load_from(data);
  std::simd_mask<float> positive = x > 0.f;
  std::simd<float>  l = std::log(std::simd_select(positive, x, 1.f));
  return std::simd_select(positive, l, x);
}
\end{lstlisting}


\section{Wording}\label{sec:wording}

Instructions to the editor:

In 29.10 Data-parallel types [simd],
\begin{enumerate}
  \item
    for every occurrence of “\code{namespace std \{}” add “\code{namespace std\wgAdd{::datapar} \{}”;

  \item
    rename all functions, types (class and alias templates) except
    \code{simd_mask}, variables (constants), and variable templates by applying
    \code{s/\backslash<simd_//} (names starting with \code{simd_} have that
    \code{simd_} prefix removed, but \code{simd_mask} is \emph{not} renamed to
  \code{mask});

  \item rename every occurrence of \code{rebind_simd} to \code{rebind};
  \item rename every occurrence of \code{resize_simd} to \code{resize};
  \item rename every occurrence of \code{rebind_simd_t} to \code{rebind_t}; and
  \item rename every occurrence of \code{resize_simd_t} to \code{resize_t}.
\end{enumerate}

Add to the end of 29.10.3 [simd.syn]
\begin{wgText}[29.10.3 Header \code{<simd>} synopsis {[simd.syn]}]
\begin{codeblock}
  template<@\exposconcept{math-floating-point}@ V>
    @\exposid{deduced-simd-t}@<V>
      sph_neumann(const rebind_simd_t<unsigned, @\exposid{deduced-simd-t}@<V>>& n, const V& x);
}

@\wgAdd{namespace std \{}@
  @\wgAdd{// [simd.alg]}@
  @\wgAdd{using datapar::min;}@
  @\wgAdd{using datapar::max;}@
  @\wgAdd{using datapar::minmax;}@
  @\wgAdd{using datapar::clamp;}@

  @\wgAdd{// [simd.math]}@
  @\wgAdd{using datapar::acos;}@
  @\wgAdd{using datapar::asin;}@
  @\wgAdd{using datapar::atan;}@
  @\wgAdd{using datapar::atan2;}@
  @\wgAdd{using datapar::cos;}@
  @\wgAdd{using datapar::sin;}@
  @\wgAdd{using datapar::tan;}@
  @\wgAdd{using datapar::acosh;}@
  @\wgAdd{using datapar::asinh;}@
  @\wgAdd{using datapar::atanh;}@
  @\wgAdd{using datapar::cosh;}@
  @\wgAdd{using datapar::sinh;}@
  @\wgAdd{using datapar::tanh;}@
  @\wgAdd{using datapar::exp;}@
  @\wgAdd{using datapar::exp2;}@
  @\wgAdd{using datapar::expm1;}@
  @\wgAdd{using datapar::frexp;}@
  @\wgAdd{using datapar::ilogb;}@
  @\wgAdd{using datapar::ldexp;}@
  @\wgAdd{using datapar::log;}@
  @\wgAdd{using datapar::log10;}@
  @\wgAdd{using datapar::log1p;}@
  @\wgAdd{using datapar::log2;}@
  @\wgAdd{using datapar::logb;}@
  @\wgAdd{using datapar::modf;}@
  @\wgAdd{using datapar::scalbn;}@
  @\wgAdd{using datapar::scalbln;}@
  @\wgAdd{using datapar::cbrt;}@
  @\wgAdd{using datapar::abs;}@
  @\wgAdd{using datapar::abs;}@
  @\wgAdd{using datapar::fabs;}@
  @\wgAdd{using datapar::hypot;}@
  @\wgAdd{using datapar::pow;}@
  @\wgAdd{using datapar::sqrt;}@
  @\wgAdd{using datapar::erf;}@
  @\wgAdd{using datapar::erfc;}@
  @\wgAdd{using datapar::lgamma;}@
  @\wgAdd{using datapar::tgamma;}@
  @\wgAdd{using datapar::ceil;}@
  @\wgAdd{using datapar::floor;}@
  @\wgAdd{using datapar::nearbyint;}@
  @\wgAdd{using datapar::rint;}@
  @\wgAdd{using datapar::lrint;}@
  @\wgAdd{using datapar::llrint;}@
  @\wgAdd{using datapar::round;}@
  @\wgAdd{using datapar::lround;}@
  @\wgAdd{using datapar::llround;}@
  @\wgAdd{using datapar::trunc;}@
  @\wgAdd{using datapar::fmod;}@
  @\wgAdd{using datapar::remainder;}@
  @\wgAdd{using datapar::remquo;}@
  @\wgAdd{using datapar::copysign;}@
  @\wgAdd{using datapar::nextafter;}@
  @\wgAdd{using datapar::fdim;}@
  @\wgAdd{using datapar::fmax;}@
  @\wgAdd{using datapar::fmin;}@
  @\wgAdd{using datapar::fma;}@
  @\wgAdd{using datapar::lerp;}@
  @\wgAdd{using datapar::fpclassify;}@
  @\wgAdd{using datapar::isfinite;}@
  @\wgAdd{using datapar::isinf;}@
  @\wgAdd{using datapar::isnan;}@
  @\wgAdd{using datapar::isnormal;}@
  @\wgAdd{using datapar::signbit;}@
  @\wgAdd{using datapar::isgreater;}@
  @\wgAdd{using datapar::isgreaterequal;}@
  @\wgAdd{using datapar::isless;}@
  @\wgAdd{using datapar::islessequal;}@
  @\wgAdd{using datapar::islessgreater;}@
  @\wgAdd{using datapar::isunordered;}@
  @\wgAdd{using datapar::assoc_laguerre;}@
  @\wgAdd{using datapar::assoc_legendre;}@
  @\wgAdd{using datapar::beta;}@
  @\wgAdd{using datapar::comp_ellint_1;}@
  @\wgAdd{using datapar::comp_ellint_2;}@
  @\wgAdd{using datapar::comp_ellint_3;}@
  @\wgAdd{using datapar::cyl_bessel_i;}@
  @\wgAdd{using datapar::cyl_bessel_j;}@
  @\wgAdd{using datapar::cyl_bessel_k;}@
  @\wgAdd{using datapar::cyl_neumann;}@
  @\wgAdd{using datapar::ellint_1;}@
  @\wgAdd{using datapar::ellint_2;}@
  @\wgAdd{using datapar::ellint_3;}@
  @\wgAdd{using datapar::expint;}@
  @\wgAdd{using datapar::hermite;}@
  @\wgAdd{using datapar::laguerre;}@
  @\wgAdd{using datapar::legendre;}@
  @\wgAdd{using datapar::riemann_zeta;}@
  @\wgAdd{using datapar::sph_bessel;}@
  @\wgAdd{using datapar::sph_legendre;}@
  @\wgAdd{using datapar::sph_neumann;}@

  @\wgAdd{// [simd.bit]}@
  @\wgAdd{using datapar::byteswap;}@
  @\wgAdd{using datapar::bit_ceil;}@
  @\wgAdd{using datapar::bit_floor;}@
  @\wgAdd{using datapar::has_single_bit;}@
  @\wgAdd{using datapar::rotl;}@
  @\wgAdd{using datapar::rotr;}@
  @\wgAdd{using datapar::bit_width;}@
  @\wgAdd{using datapar::countl_zero;}@
  @\wgAdd{using datapar::countl_one;}@
  @\wgAdd{using datapar::countr_zero;}@
  @\wgAdd{using datapar::countr_one;}@
  @\wgAdd{using datapar::popcount;}@

  @\wgAdd{// [simd.complex]}@
  @\wgAdd{using datapar::real;}@
  @\wgAdd{using datapar::imag;}@
  @\wgAdd{using datapar::arg;}@
  @\wgAdd{using datapar::norm;}@
  @\wgAdd{using datapar::conj;}@
  @\wgAdd{using datapar::proj;}@
  @\wgAdd{using datapar::polar;}@
@\wgAdd{\}}@
\end{codeblock}
\end{wgText}

\end{document}
% vim: sw=2 sts=2 ai et tw=0
