\newcommand\wgTitle{Redefining \emph{narrow contract}}
\newcommand\wgName{Matthias Kretz <m.kretz@gsi.de>}
\newcommand\wgDocumentNumber{P4011R0}
\newcommand\wgGroup{EWG}
%\newcommand\wgTarget{\CC{}26}
%\newcommand\wgAcknowledgements{Thanks to Barry Revzin for helpful feedback.}

\usepackage{mymacros}
\usepackage{wg21}
\setcounter{tocdepth}{2} % show sections and subsections in TOC
\hypersetup{bookmarksdepth=5}
\usepackage{changelog}
\usepackage{underscore}
\usepackage{multirow}

\addbibresource{extra.bib}

\newcommand\wglink[1]{\href{https://wg21.link/#1}{#1}}

\renewcommand{\lst}[1]{Listing~\ref{#1}}
\renewcommand{\sect}[1]{Section~\ref{#1}}
\renewcommand{\ttref}[1]{Tony~Table~\ref{#1}}
\renewcommand{\tabref}[1]{Table~\ref{#1}}

\begin{document}
\selectlanguage{american}
\begin{wgTitlepage}
  This paper provides a simple revised definition of narrow and wide contracts.
\end{wgTitlepage}

\pagestyle{scrheadings}

\input{changelog}
\input{strawpolls}

\section{The Problem}

\textcite{N3248} defined the terms \emph{narrow} and \emph{wide contract} in terms of
\emph{undefined behavior}:

\begin{quote}
  \textbf{Wide Contracts}

  A wide contract for a function or operation does not specify any undefined behavior.
  Such a contract has no preconditions: A function with a wide contract places no additional
  constraints on its arguments, on any object state, nor on any external global state.
  Examples of functions having wide contracts would be \code{vector<T>::begin()} and
  \code{vector<T>::at(size_type)}.
  Examples of functions not having a wide contract would be \code{vector<T>::front()} and
  \code{vector<T>::operator[](size_type)}.

  \textbf{Narrow Contracts}

  A narrow contract is a contract which is not wide.
  Narrow contracts for a functions or operations result in undefined behavior when called in a
  manner that violates the documented contract.
  Such a contract specifies at least one precondition involving its arguments, object state, or
  some external global state, such as the initialization of a static object.
  Good examples of standard functions with narrow contracts are \code{vector<T>::front()} and
  \code{vector<T>::operator[](size_type)}.
\end{quote}

Let's consider a trivial function \code{f} that results in UB when its argument is negative:
\medskip\begin{lstlisting}
void f(int x) {
  if (x < 0) std::unreachable();
}
\end{lstlisting}
Clearly, according to the definition above, this function has a precondition: \code{x >= 0}.
If we now encode that precondition using P2900 syntax we get:
\medskip\begin{lstlisting}[style=Vc]
void f(int x) pre(x >= 0) {
  if (x < 0) std::unreachable();
}
\end{lstlisting}
The trouble only begins when we introduce syntax that not only encodes the precondition, but at the
same time unconditionally removes the undefined behavior:
\medskip\begin{lstlisting}[style=Vc]
void f(int x) pre!(x >= 0) {
  if (x < 0) std::unreachable();
}
\end{lstlisting}
Since calling \code{f} can never lead to UB, the function now has a \emph{wide contract} according
to the above definition.
This is the reason why some say an unconditionally enforced precondition check is not a contract
check anymore.

I believe we need to revise the definition of narrow and wide contract, and thus the implied
definition of a \emph{contract check}.

\section{A new Definition}

\textbf{Definition:}
A function has a \emph{narrow contract} if there exists at least one input (this includes
indirect/global state that affects the function) for which the call is erroneous.
Otherwise, the function has a \emph{wide contract}.

\bigskip

\noindent\textbf{Consequence:}
A human — and potentially also a tool — can tell whether a specific call to a function is a bug in
the program or not.

\bigskip

Here “erroneous” does not have to imply erroneous behavior (but see below).
This definition aims to decouple the discussion from \emph{behavior} of the function.
For some of us this new definition makes no difference, for some it hopefully does.
The goal is to help us speak a common language.

\section{A connection to erroneous behavior?}

Consider:
\medskip\begin{lstlisting}[style=Vc]
int g(int x) pre(x >= 0) {
  if (x < 0) return 0;
  else return x + INT_MIN;
}
\end{lstlisting}

Here, the function has a precondition but also a certain behavior in case the precondition is not
met.
This is what we expect of EB\@: it is still detectable as a bug in the program, but the program can
continue without invoking UB\@.


\end{document}
% vim: sw=2 sts=2 ai et tw=100
